---
title: "Statistical Inference Course Project: PART 1"
output: pdf_document
---

# 1. Simulation Exercise

```{r echo=FALSE}
# Do 1000 simulations of 40 x Exponetials
# 
# Run and plot the simulation and explain the properties of the distribution
# 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
# 2. Show how variable it is and compare it to the theoretical variance of the distribution.
# 3. Show that the distribution is approximately normal.
# 4. Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn. (This only needs to be done for the specific value of lambda). 


# Exponential Distribution
set.seed(1234)
lambda = 0.2
actualMean = 1/lambda
actualstdDev = 1/lambda
actualVariance = actualstdDev^2
n = 40
nosim  <- 1000

# -- This code below creates a 1000 x 40 matrix with 40,000 exponentially 
# -- distributed random variables. Then it takes the mean of  
# -- each row of 1000 rand numbers and makes a tall vector.

mydata <- matrix(rexp(nosim * n, lambda) , nosim) 
distbnOfMeans <- apply(mydata, 1, mean)
# Distribution is centred at
meanOfSamples <- mean(distbnOfMeans)
# Standard Deviation of Means
stdDevOfSampleMean <- sd(distbnOfMeans)

# Optional
stdDevOfSampleStd <- sd(apply(mydata, 1, sd))

# Mean Standard Deviation of Sample
meanOfSampleStd  <- mean(apply(mydata, 1, sd))
varianceOfSample <- meanOfSampleStd^2
```

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution
```{r echo=FALSE}
rbind(
      c(1, 'Theoretical Centre of distribution', actualMean)
      ,c(1, 'Centre of Simulation Distribution' , meanOfSamples)
)
```

2. Show how variable it is and compare it to the theoretical variance of the distribution.
```{r echo=FALSE}
rbind(
      c(2, 'Theoretical Variance of Distribution', actualVariance )
      ,c(2, 'Vairance of Simulation Distribution', varianceOfSample)
)
```

3. Show that the distribution is approximately normal.
See image below: This shows the distribution of means of the 1000 sample 
Simulations is approximately normally distributed


```{r echo=FALSE, warning=FALSE, message=FALSE}
# hist(distbnOfMeans)
distbnOfMeansDf <- data.frame(distbnOfMeans)

require(ggplot2)
# m <- ggplot(distbnOfMeansDf, aes(x=distbnOfMeans) )
# m + geom_histogram(colour = "darkblue", fill = "white", binwidth = 0.08) + ggtitle('Distribution of Means of Simulations') + xlab('Means of Simulations')+ ylab('Count')


m <- ggplot(distbnOfMeansDf, aes(x = distbnOfMeans))
m <- m + geom_histogram(aes(y = ..density..), binwidth = .08 ,color="black", fill="white")
m <- m + geom_density(alpha=.2, fill="red", aes(colour="Simulated"))
m <- m + stat_function(fun = dnorm, arg = list(mean = 5, sd = 5/sqrt(40)), geom='area', alpha=.2, fill="blue", aes(colour="Normal"))
m <- m + scale_colour_manual("Density", values=c( "blue", "red"))
m <- m + scale_x_continuous(breaks = 2:9) 
m <- m + scale_y_continuous(name = "Density")
m <- m + geom_vline(data=distbnOfMeansDf, aes(xintercept=meanOfSamples,  colour= "Simulated"),
                 linetype="dashed", size=1)
m <- m + geom_vline(data=distbnOfMeansDf, aes(xintercept=5,  colour= "Normal"),
                linetype="dashed", size=1)
m <- m + theme(legend.position=c(.88, .78), 
           legend.text = element_text(size = 10, colour = "black"), 
           axis.text.x = element_text(size=15),
           axis.text.y = element_text(size=15),
           axis.title=element_text(size=20,face="bold"),
           plot.title = element_text(size=20, face="bold"))
m <- m + ggtitle('Distribution of Means of Simulations') + xlab('Means of Simulations')+ ylab('Density')
m
```

4. Evaluate the coverage of the confidence interval    
```{r echo=FALSE}
CI <- meanOfSamples + c(-1, 1) * 1.96 * meanOfSampleStd / sqrt(length(distbnOfMeans)) 
```
__Confidence Interval__ 
```{r echo=FALSE}
CI
```
__Confidence Interval Coverage__ 
```{r echo=FALSE}
# CI = 4.702276 5.302208
pnorm(5.302208, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE) - pnorm(4.702276, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE)
# 0.04942181 ~ 5%


## Tips
# Mean of each of 1000 simulations 
# sd of each of 1000 simulations

```

