---
title: "Statistical Inference Course Project"
output: pdf_document
---

# 1. Simulation Exercise

```{r echo=FALSE}
# Do 1000 simulations of 40 x Exponetials
# 
# Run and plot the simulation and explain the properties of the distribution
# 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
# 2. Show how variable it is and compare it to the theoretical variance of the distribution.
# 3. Show that the distribution is approximately normal.
# 4. Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn. (This only needs to be done for the specific value of lambda). 


# Exponential Distribution
set.seed(1234)
lambda = 0.2
actualMean = 1/lambda
actualstdDev = 1/lambda
actualVariance = actualstdDev^2
n = 40
nosim  <- 1000

# -- This code below creates a 1000 x 40 matrix with 40,000 exponentially 
# -- distributed random variables. Then it takes the mean of  
# -- each row of 1000 rand numbers and makes a tall vector.

mydata <- matrix(rexp(nosim * n, lambda) , nosim) 
distbnOfMeans <- apply(mydata, 1, mean)
# Distribution is centred at
meanOfSamples <- mean(distbnOfMeans)
# Standard Deviation of Means
stdDevOfSampleMean <- sd(distbnOfMeans)

# Optional
stdDevOfSampleStd <- sd(apply(mydata, 1, sd))

# Mean Standard Deviation of Sample
meanOfSampleStd  <- mean(apply(mydata, 1, sd))
varianceOfSample <- meanOfSampleStd^2
```

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution
```{r echo=FALSE}
rbind(
      c(1, 'Theoretical Centre of distribution', actualMean)
      ,c(1, 'Centre of Simulation Distribution' , meanOfSamples)
)
```

2. Show how variable it is and compare it to the theoretical variance of the distribution.
```{r echo=FALSE}
rbind(
      c(2, 'Theoretical Variance of Distribution', actualVariance )
      ,c(2, 'Vairance of Simulation Distribution', varianceOfSample)
)
```

3. Show that the distribution is approximately normal.
See image below: This shows the distribution of means of the 1000 sample 
Simulations is approximately normally distributed


```{r echo=FALSE, warning=FALSE, message=FALSE}
# hist(distbnOfMeans)
distbnOfMeansDf <- data.frame(distbnOfMeans)

require(ggplot2)
# m <- ggplot(distbnOfMeansDf, aes(x=distbnOfMeans) )
# m + geom_histogram(colour = "darkblue", fill = "white", binwidth = 0.08) + ggtitle('Distribution of Means of Simulations') + xlab('Means of Simulations')+ ylab('Count')


m <- ggplot(distbnOfMeansDf, aes(x = distbnOfMeans))
m <- m + geom_histogram(aes(y = ..density..), binwidth = .08 ,color="black", fill="white")
m <- m + geom_density(alpha=.2, fill="red", aes(colour="Simulated"))
m <- m + stat_function(fun = dnorm, arg = list(mean = 5, sd = 5/sqrt(40)), geom='area', alpha=.2, fill="blue", aes(colour="Normal"))
m <- m + scale_colour_manual("Density", values=c( "blue", "red"))
m <- m + scale_x_continuous(breaks = 2:9) 
m <- m + scale_y_continuous(name = "Density")
m <- m + geom_vline(data=distbnOfMeansDf, aes(xintercept=meanOfSamples,  colour= "Simulated"),
                 linetype="dashed", size=1)
m <- m + geom_vline(data=distbnOfMeansDf, aes(xintercept=5,  colour= "Normal"),
                linetype="dashed", size=1)
m <- m + theme(legend.position=c(.88, .78), 
           legend.text = element_text(size = 10, colour = "black"), 
           axis.text.x = element_text(size=15),
           axis.text.y = element_text(size=15),
           axis.title=element_text(size=20,face="bold"),
           plot.title = element_text(size=20, face="bold"))
m <- m + ggtitle('Distribution of Means of Simulations') + xlab('Means of Simulations')+ ylab('Density')
m
```

4. Evaluate the coverage of the confidence interval    
```{r echo=FALSE}
CI <- meanOfSamples + c(-1, 1) * 1.96 * meanOfSampleStd / sqrt(length(distbnOfMeans)) 
```
__Confidence Interval__ 
```{r echo=FALSE}
CI
```
__Confidence Interval Coverage__ 
```{r echo=FALSE}
# CI = 4.702276 5.302208
pnorm(5.302208, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE) - pnorm(4.702276, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE)
# 0.04942181 ~ 5%


## Tips
# Mean of each of 1000 simulations 
# sd of each of 1000 simulations

```




# 2. Basic Inferential Data Analysis
```{r echo=FALSE}
# 1. Load the ToothGrowth data and perform some basic exploratory data analyses 
# 2. Provide a basic summary of the data.
# 3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the  techniques from class even if theres other approaches worth considering)
# 4. State your conclusions and the assumptions needed for your conclusions. 

```

## 1. Load the Tooth Growth Dataset & Basic Exploratory Analysis
```{r echo=FALSE, warning=FALSE, message=FALSE}
require(data.table)
library(datasets); data(ToothGrowth);
dt <- data.table(ToothGrowth)
```

Type of Supp  | Number Trials
:-------------: | :-------------:
VC  | 30
OJ  | 30

Dose (milligrams)  | Number Trials
:-------------: | :-------------:
0.5 | 20
1.0 | 20
2.0 | 20

## 2. Basic Summary of Data
```{r echo=FALSE}
require(ggplot2)
##summary(ToothGrowth)
```

### Supplements and Doses Summary Table
```{r echo=FALSE}
dt[, list(meanLength=mean(len), sdLength=sd(len)), by = c('supp', 'dose')]

#    supp dose meanLength sdLength
# 1:   VC  0.5       7.98 2.746634
# 2:   VC  1.0      16.77 2.515309
# 3:   VC  2.0      26.14 4.797731
# 4:   OJ  0.5      13.23 4.459709
# 5:   OJ  1.0      22.70 3.910953
# 6:   OJ  2.0      26.06 2.655058
```
### Supplements Summary Table
```{r echo=FALSE}
dt[, list(meanLength=mean(len), sdLength=sd(len)), by = supp]
#    supp meanLength sdLength
# 1:   VC   16.96333 8.266029
# 2:   OJ   20.66333 6.605561
```
### Doses Summary Table
```{r echo=FALSE}
dt[, list(meanLength=mean(len), sdLength=sd(len)), by = dose]
#    dose meanLength sdLength
# 1:  0.5     10.605 4.499763
# 2:  1.0     19.735 4.415436
# 3:  2.0     26.100 3.774150
```
### Graphical Summary of Supplements and Doses
```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(dt, aes(len) ) + geom_histogram() + facet_wrap(supp ~ dose)
```

## 3. Use Confidence Intervals and Hypothesis Tests to compare tooth growth by supplement and dose

### a) Confidence Intervals

__Comparing by Supp__  
Independent t confidence interval for OJ - VC
```{r echo=FALSE}
###############
# Paired
###############
# Y
n_OJ = 30
X_OJ = 20.66333
s_OJ = 6.605561

# X
n_VC = 30
X_VC = 16.96333
s_VC = 8.266029

alpha = 0.05

degfree = n_OJ + n_VC - 2

Sp = sqrt( ( (n_VC - 1) * s_VC^2 + (n_OJ - 1) * s_OJ^2 ) /  degfree )

# t for 95%
round(X_OJ - X_VC + c(-1,1) * qt(1 - alpha/2, degfree) * Sp * sqrt(1/n_VC + 1/n_OJ), 2)

# CI = -0.17,  7.57
# Most of the time the OJ out-performs the VC with 95% confidence

```
Most of the time the OJ out performs the VC with 95% confidence


__Comparing by Dose__
```{r echo=FALSE}
# dt[, list(meanLength=mean(len), sdLength=sd(len)), by = dose]
#    dose meanLength sdLength
# 1:  0.5     10.605 4.499763
# 2:  1.0     19.735 4.415436
# 3:  2.0     26.100 3.774150

# X
n_05 = 20
X_05 = 10.605
s_05 = 4.499763

# Y
n_10 = 20
X_10 = 19.735
s_10 = 4.415436

# Z
n_20 = 20
X_20 = 26.100
s_20 = 3.774150

# 95% CI
alpha = 0.05

## Compare 0.5 vs 1.0 
degfree = n_10 + n_05 - 2

Sp = sqrt( ( (n_05 - 1) * s_05^2 + (n_10 - 1) * s_10^2 ) /  degfree )
```
### Independent t confidence interval comparing doses 0.5 vs 1.0 milligrams
```{r echo=FALSE}
# t for 95%
round(X_10 - X_05 + c(-1,1) * qt(1 - alpha/2, degfree) * Sp * sqrt(1/n_05 + 1/n_10), 2)

# CI = 6.28 11.98
# Dose of 1.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence
```
Dose of 1.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence

```{r echo=FALSE}

## Compare 0.5 vs 2.0 
degfree = n_20 + n_05 - 2

Sp = sqrt( ( (n_05 - 1) * s_05^2 + (n_20 - 1) * s_20^2 ) /  degfree )
```
### Independent t confidence interval comparing doses 0.5 vs 2.0 milligrams
```{r echo=FALSE}
# t for 95%
round(X_20 - X_05 + c(-1,1) * qt(1 - alpha/2, degfree) * Sp * sqrt(1/n_05 + 1/n_20), 2)

# CI = 12.84 18.15
# Dose of 2.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence
```
Dose of 2.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence

```{r echo=FALSE}
## Compare 1.0 vs 2.0 
degfree = n_20 + n_10 - 2

Sp = sqrt( ( (n_10 - 1) * s_10^2 + (n_20 - 1) * s_20^2 ) /  degfree )
```
### Independent t confidence interval comparing doses 2.0 vs 1.0 milligrams
```{r echo=FALSE}
# t for 95%
round(X_20 - X_10 + c(-1,1) * qt(1 - alpha/2, degfree) * Sp * sqrt(1/n_10 + 1/n_20), 2)

# CI = 3.74 8.99
# Dose of 2.0 always out performs does of 1.0 in stimulating tooth growth with 95% confidence
```
Dose of 2.0 always out performs does of 1.0 in stimulating tooth growth with 95% confidence


### b) Hypothesis Test
__Comparing by Supp__
```{r echo=FALSE}
## H0 = VC is a better supplement that OJ
## Ha = OJ is better
## If p-value > 0.05 then reject H0. Therefore OJ is a better supplement than VC


t.test(dt[supp=='OJ']$len , dt[supp=='VC']$len, paired = FALSE, var.equal = FALSE)
# Alternative notation
# t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data=dt)

# Welch Two Sample t-test
# 
# data:  dt[supp == "OJ"]$len and dt[supp == "VC"]$len
# t = 1.9153, df = 55.309, p-value = 0.06063
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -0.1710156  7.5710156
# sample estimates:
# mean of x mean of y 
#  20.66333  16.96333 

dt[, list(mean=mean(len), sd=sd(len))]

X = 16.9333
mu = 18.81333
s = 7.649315
n = nrow(dt)
TS = (X - mu) / (s / sqrt(n))

#let 
alpha = 0.05 
Z_alpha = qnorm(alpha) 

# Reject Null Hypothesis when
if ( TS <= Z_alpha ) 'Reject H0' else 'Fail to Reject H0'

# Question 
# How do you know what the mean of the population is?
# E.g. in this project we only know the mean of the sample from the Tooth Growth data




```

__Comparing by Dose__
```{r echo=FALSE}
## 0.5 vs 1.0
## H0 = dose 1.0
## Ha = dose 0.5
## if p-value < 0.05 then fail to reject H0. Therefore dose 1.0 is better than 0.5
t.test(dt[dose==0.5]$len , dt[dose==1.0]$len, paired = FALSE, var.equal = FALSE)
#       Welch Two Sample t-test
# 
# data:  dt[dose == 0.5]$len and dt[dose == 1]$len
# t = -6.4766, df = 37.986, p-value = 1.268e-07
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -11.983781  -6.276219
# sample estimates:
# mean of x mean of y 
#    10.605    19.735 

## 0.5 vs 2.0
t.test(dt[dose==0.5]$len , dt[dose==2.0]$len, paired = FALSE, var.equal = FALSE)

#       Welch Two Sample t-test
# 
# data:  dt[dose == 0.5]$len and dt[dose == 2]$len
# t = -11.799, df = 36.883, p-value = 4.398e-14
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -18.15617 -12.83383
# sample estimates:
# mean of x mean of y 
#    10.605    26.100 

## 1.5 vs 2.0
t.test(dt[dose==1.0]$len , dt[dose==2.0]$len, paired = FALSE, var.equal = FALSE)
#       Welch Two Sample t-test
# 
# data:  dt[dose == 1]$len and dt[dose == 2]$len
# t = -4.9005, df = 37.101, p-value = 1.906e-05
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -8.996481 -3.733519
# sample estimates:
# mean of x mean of y 
#    19.735    26.100 


```

## 4. Conclusions and Assumptions 
### Conclusions
1. Most of the time the OJ out performs the VC with 95% confidence
2. Dose of 1.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence
3. Dose of 2.0 always out performs does of 0.5 in stimulating tooth growth with 95% confidence
4. Dose of 2.0 always out performs does of 1.0 in stimulating tooth growth with 95% confidence

### Assumptions
1. T distribution is used as data may not be normally distributed
2. Assume unequal variances for t distribution confidence interval
3. Assume the Central Limit Theorm
4. n must be large enough to be statistically signifcant
5. Since n is small for each set of tests use Gossett's T test
6. Assuming a constant variance between groups of Guinea Pigs receiving difference amounts of treatment and different supplements 















