---
title: "Statistical Inference Course Project"
author: "Andrew Szwec"
date: "23 August 2014"
output: pdf_document
---

# 1. Simulation Exercise

Do 1000 simulations of 40 x Exponetials

Run and plot the simulation and explain the properties of the distribution
1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.
4. Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn√. (This only needs to be done for the specific value of lambda). 

```{r}
# Exponential Distribution
set.seed(1234)
lambda = 0.2
actualMean = 1/lambda
actualstdDev = 1/lambda
actualVariance = actualstdDev^2
n = 40
nosim  <- 1000

# -- This code below creates a 1000 x 40 matrix with 40,000 exponentially 
# -- distributed random variables. Then it takes the mean of  
# -- each row of 1000 rand numbers and makes a tall vector.

mydata <- matrix(rexp(nosim * n, lambda) , nosim) 
distbnOfMeans <- apply(mydata, 1, mean)
# Distribution is centred at
meanOfSamples <- mean(distbnOfMeans)
# Standard Deviation of Means
stdDevOfSampleMean <- sd(distbnOfMeans)

# Optional
stdDevOfSampleStd <- sd(apply(mydata, 1, sd))

# Mean Standard Deviation of Sample
meanOfSampleStd  <- mean(apply(mydata, 1, sd))
varianceOfSample <- meanOfSampleStd^2

rbind(
      c(1, 'Theoretical Centre of distribution', actualMean)
      ,c(1, 'Centre of Simulation Distribution' , meanOfSamples)
      ,c(2, 'Theoretical Variance of Distribution', actualVariance )
      ,c(2, 'Vairance of Simulation Distribution', varianceOfSample)
)

# See image below: This shows the distribution of means of the 1000 sample 
# Simulations is approximately normally distributed
hist(distbnOfMeans)

distbnOfMeansDf <- data.frame(distbnOfMeans)
m <- ggplot(distbnOfMeansDf, aes(x=distbnOfMeans) )
m + geom_histogram(colour = "darkblue", fill = "white", binwidth = 0.5) + ggtitle('Distribution of Means of Simulations') + xlab('Means of Simulations')+ ylab('Count')


# Confidence Interval
# Std Dev = 1/lambda, Mean = 1/lambda
# X_sample +/- 1.96 * S / sqrt(n)

CI <- meanOfSamples + c(-1, 1) * 1.96 * meanOfSampleStd/ sqrt(length(distbnOfMeans)) 
# CI = 4.702276 5.302208
pnorm(5.302208, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE) - pnorm(4.702276, mean = meanOfSamples, sd = meanOfSampleStd, lower.tail = TRUE)
# 0.04942181 ~ 5%

 


```




# 2. Basic Inferential Data Analysis

Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.
Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions. 


## 1. Load the Tooth Growth Dataset & Basic Exploratory Analysis
```{r}
require(data.table)
library(datasets); data(ToothGrowth);
dt <- data.table(ToothGrowth)
# Types of Supplement 
unique(dt$supp)
# Doses Used in Trial
unique(dt$dose)
```

## 2. Basic Summary of Data
```{r}
summary(ToothGrowth)
dt[, list(meanLength=mean(len), sdLength=sd(len)), by = c('supp', 'dose')]

#    supp dose meanLength sdLength
# 1:   VC  0.5       7.98 2.746634
# 2:   VC  1.0      16.77 2.515309
# 3:   VC  2.0      26.14 4.797731
# 4:   OJ  0.5      13.23 4.459709
# 5:   OJ  1.0      22.70 3.910953
# 6:   OJ  2.0      26.06 2.655058

ggplot(dt, aes(len) ) + geom_histogram() + facet_wrap(supp ~ dose)
```

## 3. Use Confidence Intervals and Hypothesis Tests to compare tooth growth by supplement and dose

### a) Confidence Intervals
```{r}


```

### b) Hypothesis Test
```{r}



```

## 4. Conclusions and Assumptions 
1.
2.
3.
4.





















